---
title: "Capstone Analysis Queries"
author: "Maddi Klepps"
date: "2026-03-22"
output:
  pdf_document: default
  html_document: default
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# **Bring in Cleaned & Transformed Dataset**

```{r}
Dataset <- read.csv("Cleaned_Transformed_Dataset_Klepps.csv")
```


# **Import Needed Libraries**

```{r, message = FALSE, warning = FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
```


# **Exploratory Analysis**


## **Counts**

```{r}
Dataset |>
  summarize(unique_count=n_distinct(author))
```

```{r}
Dataset |>
  summarize(unique_count=n_distinct(title))
```

```{r}
Dataset |>
  summarize(unique_count=n_distinct(genre))
```

## **Authors**

### How many times each Author appears (output hidden due to hundreds of lines)
```{r, results = "hide"}
author_counts <- Dataset |>
  count(author, name = "count")

author_counts
```

### Most popular Authors (output hidden due to hundreds of lines)
```{r,  results = "hide"}
author_rank <- Dataset |>
  count(author, name = "count") |>
  arrange(desc(count))

author_rank
```



## **Titles**

### How many times each Title appears
### *slightly redundant to above query - verifies each title only appears once?* 
```{r}
title_counts <- Dataset |>
  count(title, name = "count")
```

### **Title Ratings**
```{r}
title_ratings <- Dataset |>
  arrange(desc(rating)) |>
  mutate(order_rating = row_number())

result_ratings <- title_ratings |>
  select(title, author, rating)
```



## **Genres**

### How many time each Genre appears
```{r}
genre_counts <- Dataset |>
  count(genre, name = "count") |>
  arrange(desc(count))
```

### How many time each Subgenre appears
```{r}
subgenre_counts <- Dataset |>
  count(subgenre, name = "count") |>
  arrange(desc(count))
```

### Genres & associated Subgenres
```{r}
genres_subgenres_nested <- Dataset |>
  group_by(genre) |>
  summarize(subgenres = list(unique(subgenre)))
```

```{r, results = "hide", warning = FALSE}
split(unique(Dataset$subgenre), Dataset$genre)
```


# **Refined Analysis & Querying Setup**


## **Packages**

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(scales)
library(stringr)
```


## **Theme Creation**

```{r}
theme_sage_purple <- function() {
  theme_bw() +
  theme(
    # Axis Text
    axis.text.y = element_text(size = 9, lineheight = 1.1),
    axis.text.x = element_text(size = 10),
    
    # Titles
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5, margin = margin(b = 10)),
    axis.title.y = element_text(size = 12, face = "bold"),
    axis.title.x = element_text(size = 12, face = "bold"),
    
    # Background
    panel.background = element_rect(fill = "#DDE8D8", color = NA),
    plot.background = element_rect(fill = "#DDE8D8", color = NA),
    
    # Gridlines
    panel.grid.major = element_line(color = "white"),
    panel.grid.minor = element_line(color = "white"),
    
    #Margins
    plot.margin = margin(10, 40, 10, 10)
  )
}
```



# **Refined Queries**


## **1. Which genres generate the most total revenue? (using number of reviews multiplied by final selling price)**
```{r}
Dataset <- Dataset |>
  mutate(revenue = final_price * reviews_count)

genre_revenue <- Dataset |>
  group_by(genre) |>
  summarize(total_revenue = sum(revenue, na.rm = TRUE)) |>
  arrange(desc(total_revenue))
```

```{r fig.width=14, fig.height=10}
ggplot(genre_revenue, aes(x = reorder(str_wrap(genre, width = 25), total_revenue), y = total_revenue)) +
  geom_col(fill = "#796878") +
  geom_text(
    aes(label = paste0(round(total_revenue/ 1e6, 1), "M")),
    hjust = -0.1,
    size = 3
  ) +
  coord_flip() +
  scale_y_continuous(
    labels = label_number(scale = 1e-6, suffix = "M"),
    expand = expansion(mult =c(0, 0.1))
  ) +
  labs(
    title = "Total Revenue by Genre",
    x = "Genre",
    y = "Total Revenue (Millions USD)"
  ) +
  theme_sage_purple() +
  guides(fill = "none")
```
  

## **2. Which sub-genres generate the most total revenue? (using number of reviews multiplied by final selling price)**
```{r}
Dataset <- Dataset |>
  mutate(revenue2 = final_price * reviews_count)

subgenre_revenue <- Dataset |>
  group_by(subgenre) |>
  summarize(total_revenue = sum(revenue2, na.rm = TRUE)) |>
  arrange(desc(total_revenue)) |>
  slice_head(n = 10)
``` 

```{r fig.width=14, fig.height=6}
ggplot(subgenre_revenue, aes(x = reorder(str_wrap(subgenre, width = 25), total_revenue), y = total_revenue)) +
  geom_col(fill = "#796878") +
  geom_text(
    aes(label = paste0(round(total_revenue/ 1e6, 1), "M")),
    hjust = -0.1,
    size = 3
  ) +
  coord_flip() +
  scale_y_continuous(
    labels = label_number(scale = 1e-6, suffix = "M"),
    expand = expansion(mult =c(0, 0.1))
    ) +
  labs(
    title = "Total Revenue by Subgenre",
    x = "Subgenre",
    y = "Total Revenue (Millions USD)"
  ) +
  theme_sage_purple() +
  guides(fill = "none")
```


## ***3. Which individual bestselling books drive the most revenue? **

```{r fig.width = 14, fig.heigh = 10}
top_books <- Dataset |>
  mutate(revenue = final_price * reviews_count) |>
  arrange(desc(revenue)) |>
  slice_head(n = 10)

print(top_books |>
        select(title, author, genre, subgenre, revenue))

ggplot(top_books, aes(x = reorder(str_wrap(title, width = 30), revenue), y = revenue)) +
  geom_col(fill = "#796878") +
  geom_text(
    aes(label = paste0(round(revenue / 1e6, 1), "M")),
    hjust = -0.1,
    size = 3
  ) +  
  coord_flip() +
  scale_y_continuous(
    labels = label_number(scale = 1e-6, suffix = "M"),
    expand = expansion(mult = c(0, 0.1))
  ) +
  labs(
    title = "Top 10 Books by Revenue",
    x = "Book Title",
    y = "Total Revenue (Millions USD)"
  ) +
  theme_sage_purple() +
  guides(fill = "none")
```


## **4. Do highly rated books mean stronger customer engagement? (relationship between rating & review)**

```{r fig.width=14, fig.height=8}

Dataset_filtered <- Dataset |>
  filter(reviews_count < 80000) #remove outlier values to better show trend

ggplot(Dataset_filtered, aes(x = rating, y = reviews_count)) +
  geom_jitter(width = 0.02, alpha = 0.35, size = 1.5, color = "#796B87") +
  geom_smooth(method = "lm", color = "mediumpurple4", se = FALSE, linewidth = 1.2) +
  scale_y_continuous(labels = label_number(scale = 1e-3, suffix = "K")) +
  labs(
    title = "Relationship Between Rating & Customer Engagement",
    subtitle = "Customer Engagement is measured by review count",
    caption = "Books with review counts over 80,000 were excluded to improve visualization & clearly display trend.",
    x = "Average Rating (out of 5.0)",
    y = "Number of Reviews"
  ) +
  theme_sage_purple() +
  theme(plot.subtitle = element_text(hjust = 0.5, size = 11))
```

## **alternative Boxplot to display info in another way**

```{r fig.width=14, fig.height=8}
Dataset_filtered <- Dataset |>
  filter(reviews_count < 80000) |>
  mutate(rating_group = factor(round(rating, 1)))

ggplot(Dataset_filtered, aes(x = rating_group, y = reviews_count)) +
  geom_boxplot(fill = "#796B87", color = "#4B3FBF", alpha = 0.8, outlier.alpha = 0.3) +
  geom_jitter(width = 0.15, alpha = 0.15, size = 1, color = "#4B3FBF") +
  scale_y_continuous(labels = label_number(scale = 1e-3, suffix = "K")) +
  labs(
    title = "Customer Engagement by Rating Group",
    subtitle = "Customer engagement is measured by review count",
    x = "Average Rating (Grouped)",
    y = "Number of Reviews",
    caption = "Books with review counts above 80,000 were excluded to improve visualization clarity."
  ) +
  theme_sage_purple() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 11),
    panel.grid.minor = element_blank()
  )
```



## **5. Which books should be prioritized moving forward? (revenue & engagement)**

```{r fig.width=14, fig.height=10}
books_priority <- Dataset |>
  mutate(
    revenue = final_price * reviews_count,
    revenue_z = as.numeric(scale(revenue)),
    engagement_z = as.numeric(scale(reviews_count)),
    priority_score = 0.6 * revenue_z + 0.4 * engagement_z
  ) |>
  arrange(desc(priority_score))

head(books_priority |>
       select(title, author, genre, subgenre, revenue, reviews_count, priority_score), 10)

top_priority <- books_priority |>
  slice_head(n = 10)

ggplot(top_priority, aes(x = reorder(str_wrap(title, width = 35), priority_score), y = priority_score)) +
  geom_col(fill = "#796B87") +
  geom_text(
    aes(label = paste0(round(revenue / 1e6, 1), "M")),
    hjust = 1.2,
    color = "white",
    size = 3
  ) +
  coord_flip() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))
  ) +
  labs(
    title = "Top 10 Products to Prioritize Moving Forward",
    subtitle = "Based on a combined score of revenue (60%) and customer engagement (40%).",
    x = "Book Title",
    y = "Priority Score"
  ) +
  theme_sage_purple() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 11),
    panel.grid.minor = element_blank()
  ) +
  guides(fill = "none")
```

## **Experimental to show contributions to Priority Score**

```{r fig.width=16, fig.height=9}  
top_priority_long <- top_priority |>
  select(title, priority_score, revenue_z, engagement_z) |>
  mutate(title = str_wrap(title, width = 30))

p1 <- top_priority_long |>
  select(title, priority_score) |>
  ggplot(aes(x = reorder(title, priority_score), y = priority_score)) +
  geom_col(fill = "#796B87") +
  geom_text(aes(label = round(priority_score, 2)), hjust = -0.1, size = 3) +
  coord_flip() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(
    title = "Priority Score Ranking",
    x = "Book Title",
    y = "Priority Score"
  ) +
  theme_sage_purple() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    panel.grid.minor = element_blank()
  )

p2 <- top_priority_long |>
  pivot_longer(
    cols = c(revenue_z, engagement_z),
    names_to = "component",
    values_to = "value"
  ) |>
  mutate(
    component = recode(component,
                       revenue_z = "Revenue",
                       engagement_z = "Engagement")
  ) |>
  ggplot(aes(x = reorder(title, value), y = value, fill = component)) +
  geom_col(position = "dodge") +
  coord_flip() +
  scale_fill_manual(values = c("Revenue" = "#796B87", "Engagement" = "#4B3FBF")) +
  labs(
    title = "Revenue vs Engagement Contribution",
    x = "Book Title",
    y = "Standardized Contribution",
    fill = "Component"
  ) +
  theme_sage_purple() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    panel.grid.minor = element_blank()
  )

p2
```
































